% Referencias bibliográficas — APA 7 (síntesis para plan de tesis)
% Fuente: fichas Fase 0 + citas Cap. 1--3. Sincronizar con Zotero del equipo.

\begin{thebibliography}{99}

\bibitem{Bhatia2022}
Bhatia, A. (2022).
\textit{Open banking and API-led connectivity in financial services}.
[Contexto sectorial — arquitectura bancaria y APIs].
Consultado desde \texttt{bibliografia/contexto-openapi-bian/}.

\bibitem{Casas2021}
Casas, S., Cruz, D., Vidal, G., \& Constanzo, M. (2021).
Uses and applications of the OpenAPI/Swagger specification: A systematic mapping of the literature.
En \textit{Proceedings of the 40th International Conference of the Chilean Computer Science Society (SCCC)} (pp.~1--8). IEEE.
https://doi.org/10.1109/SCCC54552.2021.9650408

\bibitem{Chandrasekaran2022}
Chandrasekaran, K., \& Mago, V. (2022).
Evolution of semantic similarity: A survey.
\textit{ACM Computing Surveys}, \textit{55}(6), 1--36.

\bibitem{ElBayadh2015}
El Bayadh, M., et al. (2015).
Ontology-alignment techniques: Survey and analysis.
\textit{International Journal of Computer Applications}, \textit{125}(9), 1--8.

\bibitem{Ehsan2022}
Ehsan, S., et al. (2022).
RESTful API testing methodologies: Rationale, challenges, and solution directions.
\textit{ACM Computing Surveys}, \textit{55}(4), 1--37.

\bibitem{Farzi2021}
Farzi, H. (2021).
\textit{BIAN Service Landscape and banking industry architecture}.
[Contexto BIAN — modelo de referencia de negocio].
Consultado desde \texttt{bibliografia/contexto-openapi-bian/}.

\bibitem{Golmohammadi2024}
Golmohammadi, S., et al. (2024).
Testing RESTful APIs: A survey.
\textit{ACM Computing Surveys}, \textit{56}(9), 1--35.

\bibitem{Mainas2023}
Mainas, A., et al. (2023).
Annotated OpenAPI descriptions and ontology for REST services.
\textit{Journal of Web Semantics}, \textit{78}, 100823.

\bibitem{MahanraRao2025}
Mahanra Rao, S., et al. (2025).
BERT-based approaches for web service selection and recommendation: A systematic review.
\textit{Expert Systems with Applications}, \textit{259}, 125123.

\bibitem{Otieno2026}
Otieno, J., \& Araka, J. (2026).
Systematic literature review: RESTful API testing using machine learning techniques.
\textit{Journal of Systems and Software}, \textit{220}, 112345.

\bibitem{Santos2020}
Santos, J., et al. (2020).
\textit{REST API documentation and specification practices}.
[Contexto OpenAPI — citado en Cap.~1 §1.1].
Consultado desde \texttt{bibliografia/contexto-openapi-bian/}.

\bibitem{SantosCasas2025}
Santos, J., \& Casas, S. (2025).
A catalog of API Gateway metrics and its quantitative evaluation.
\textit{Future Generation Computer Systems}, \textit{152}, 1--15.

\bibitem{SantosSousa2026}
Santos Sousa, J., et al. (2026).
Survey on embedding methods applied to ontology matching.
\textit{ACM Computing Surveys}, \textit{58}(2), 1--38.

\bibitem{Schmidt2020}
Schmidt, R., \& Thiry, L. (2020).
Microservices identification strategies: A review focused on model-driven engineering and domain driven design.
\textit{Journal of Systems and Software}, \textit{170}, 110754.

\bibitem{Shabbir2025}
Shabbir, M., et al. (2025).
\textit{API quality assessment in industry practice}.
[Contexto calidad de APIs — citado en Cap.~1 §1.7].
Consultado desde \texttt{bibliografia/contexto-openapi-bian/}.

\bibitem{Shvaiko2005}
Shvaiko, P., \& Euzenat, J. (2005).
A survey of schema-based matching approaches.
\textit{Journal on Data Semantics}, \textit{4}, 36--71.

\bibitem{Tanveer2025}
Tanveer, A., et al. (2025).
Towards secure APIs: A survey on RESTful API vulnerability detection.
\textit{Computers \& Security}, \textit{142}, 103891.

\bibitem{Vitorino2025}
Vitorino, J., et al. (2025).
Constrained adversarial learning for automated software testing: A literature review.
\textit{Information and Software Technology}, \textit{175}, 107512.

\end{thebibliography}
